---
title: "§4 Dealing with Cards 🃏"
subtitle: "R Objects and Data Frames"
format:
  html:
    toc: true
    code-fold: false
    code-link: true
    highlight-style: github
    include-in-header:
      text: |
        <script src="quiz.js"></script>
        <script src="content-switch.js"></script>
filters: 
 - collapse-callout.lua
execute:
  freeze: auto
---

Welcome to your second project with R! We'll create a virtual deck of poker cards to explore more advanced R concepts. 

::: {.callout-tip}
## Learning Objectives
- 🧱 Understand different R data types - the building blocks of information
- 📊 Create and manipulate data frames - R's powerful spreadsheet-like structure
- 🃏 Build a virtual deck of cards - applying R concepts to a real-world object
- 🔍 Extract and modify data - skills crucial for data analysis
:::

# R Data Types 🧱

In R, just as in language, we have different types of information. Let's explore some common data types:

::: {.callout-note}
## Basic Data Types
- Numeric: For numbers (integers and decimals)
- Character: For text data
- Logical: For true/false values
- Factor: For categorical data
Each type serves a specific purpose in organizing and analyzing information.
:::

## Numeric
We've already seen this with our dice. It includes both integers and decimal numbers:

```{r}
my_integer <- 42
my_deci <- 3.14
class(my_integer)
class(my_deci)
```

## Character
This is used for text data, like words or sentences:

```{r}
my_name <- "Shakespeare"
my_quote <- "To be or not to be"
class(my_name)
class(my_quote)
```


## Logical
This represents true/false values:

```{r}
is_sunny <- TRUE
is_raining <- FALSE
class(is_sunny)
class(is_raining)
```


## Factor
This is used for categorical data, like genres in literature:

```{r}
genres <- factor(c("Poetry", "Prose", "Drama"))
class(genres)
```


::: {.callout-tip}
## What is the Difference Between Factor and Character?

Character and factor may seem similar, but they serve different purposes:

1. [Character]{.highlight-blue}
   - This is simply text data.
   - It can be any combination of letters, numbers, or symbols.
   - Examples: `"apple"`, `"banana"`, `"cherry123"`

2. [Factor]{.highlight-green}
   - This is for categorical data, made up of distinct levels or categories.
   - It's used when data falls into a limited number of specific groups.
   - Examples: Days of the week, book genres, or rating scales


**Key differences:**

- Characters can be any text, while factors consist of predefined levels or categories.
- Each level in a factor is assigned a unique integer behind the scenes, making factors more efficient for certain analyses.
- Factors are ideal for data with a fixed set of possible values or categories.

:::

## Learning Check 🏁 

<div id="quiz1"></div>

<script>
createQuiz({
  questions: [
    {
      text: "Which data type would you use to store a person's age?",
      options: [
        "Character",
        "Numeric",
        "Logical",
        "Factor"
      ]
    },
    {
      text: "What data type is best for storing the title of a book?",
      options: [
        "Numeric",
        "Character",
        "Logical",
        "Factor"
      ]
    },
    {
      text: "Which data type would you use to represent whether a book is fiction, poetry, or documentary?",
      options: [
        "Numeric",
        "Character",
        "Logical",
        "Factor"
      ]
    }
  ],
  answers: [1, 1, 3]
}, "quiz1");
</script>

## Hands-On Coding 💻

Try the following:

1. Create a numeric variable called `year_published` and assign it the year your favorite book was published.
2. Create a character variable called `book_title` with the title of your favorite book.
3. Create a logical variable called `is_fiction` indicating whether your favorite book is fiction (TRUE) or non-fiction (FALSE).
4. Create a factor variable called `book_genre` with a few genres (e.g., "Mystery", "Romance", "Science Fiction").

<div id="hidden-content1" style="display: none;">


```{r}
# Year published
year_published <- 1960
# Book title
book_title <- "To Kill a Mockingbird"
# Fiction or non-fiction
is_fiction <- TRUE
# Book genre
book_genre <- factor(c("Mystery", "Romance", "Science Fiction"))
```

</div>

<script>
createToggleSection('button1', 'hidden-content1');
</script>

# Data Frames 📊

A data frame is like a spreadsheet in R. It's a collection of vectors of equal length, each representing a column. This structure is perfect for our deck of cards!

Let's create a simple data frame to represent books:


```{r}
books <- data.frame(
title = c("1984", "Pride and Prejudice", "The Great Gatsby"),
author = c("Orwell", "Austen", "Fitzgerald"),
year = c(1949, 1813, 1925)
)
print(books)
str(books)
```

::: {.callout-note}
## `str()`
The `str()` function in R stands for "structure". It's a very useful tool for quickly examining the structure of any R object, especially data frames. Here's what it shows you:

1. The type of object (e.g., data frame, list, vector)
2. The number of observations (rows) and variables (columns) for data frames
3. The name and type of each variable (column)
4. A preview of the data in each variable

For example, when we use `str(books)`, it might show:
```
'data.frame': 3 obs. of 3 variables:
$ title : chr "1984" "Pride and Prejudice" "The Great Gatsby"
$ author: chr "Orwell" "Austen" "Fitzgerald"
$ year : num 1949 1813 1925
```

This tells us:
- We have a data frame
- It has 3 observations (rows) and 3 variables (columns)
- The variables are:
  - `title`: character type (text)
  - `author`: character type (text)
  - `year`: numeric type (numbers)

Using `str()` is a quick way to get an overview of your data, which is especially helpful when working with large datasets.
:::


## Accessing Data in a Data Frame

You can access specific parts of a data frame using various methods.

::: {.callout-note}
## `$` and `[row_number, column_number]`

1. **Columns**: Use `$` followed by the column name (e.g., `books$title`).
   - This gives you all the values in that column.

2. **Rows**: Use `[row_number, ]` (e.g., `books[2, ]`).
   - The comma is important! It means "all columns".
   - This gives you all the data in that specific row.

3. **Individual Cells**: Use `[row_number, column_number]` (e.g., `books[1, 2]`).
   - This gives you the value in a specific row and column.

:::

::: {.callout-tip}
## Acessing the `books` Data Frame
Data frames are essentially just spread sheet tables. Our `books` data frame looks like this:

| title               | author    | year |
|---------------------|-----------|------|
| 1984                | Orwell    | 1949 |
| Pride and Prejudice | Austen    | 1813 |
| The Great Gatsby    | Fitzgerald| 1925 |

- `books$title` would give: `1984`, `Pride and Prejudice`, `The Great Gatsby`
- `books[2, ]` would give: `Pride and Prejudice`, `Austen`, `1813`
- `books[1, 2]` would give: `Orwell`
:::


```{r}
# Get a specific column
books$title

# Get a specific row
books[2, ]

# Get a specific cell
books[1, 2]

# Get a specific cell using row number and column name
books[2, "title"] # Get the second title

# Get the record based on a specific value
books[books$author == "Austen", ] # Get all records by Austen
```

::: {.callout-note}
## Understanding Nested Functions
The line `books[books$author == "Austen", ]` might look complex, but let's break it down:

1. `books$author`: This part gets the `author` column from our books data frame.
2. `books$author == "Austen"`: This compares each author name to "Austen".
- It creates a list of TRUE/FALSE values (TRUE where the author is Austen, FALSE otherwise).
- For example, if we had 3 books and only the second was by Austen, this would give: [FALSE, TRUE, FALSE]
3. `books[...]`: This is like asking R to look inside the books data frame.
4. `books[books$author == "Austen", ]`: This combines steps 2 and 3.
- It tells R: "From the books data frame, give me all rows where the author is Austen".
- The comma at the end means "give me all columns for these rows".

Think of it like a librarian (R) searching through a catalog (the data frame):

- You ask: "Can you find all books by Austen?"
- The librarian checks each book's author (TRUE/FALSE for Austen).
- Then returns all information about the books that matched.
:::

::: {.callout-tip}
## Filtering Data Based on Conditions
This example demonstrates a powerful technique called "filtering":

1. Creating a condition: `books$author == "Austen"` creates a logical vector (TRUE/FALSE) for each row where the author is Austen.
2. Using the condition: We use this vector inside the square brackets `[ ]` to select only those rows that meet our condition.
3. Selecting columns: The comma after the condition (`books [condition, ]`) means "select all columns for these rows".

This method allows you to extract specific records based on any condition you specify. For example:

- Find all books published after 1900: books[books$year > 1900, ]
- Find all books with "War" in the title: books[grepl("War", books$title), ]

Filtering is a crucial skill in data manipulation and analysis, allowing you to focus on specific subsets of your data.
:::

## Learning Check 🏁

<div id="quiz2"></div>

<script>
createQuiz({
  questions: [
    {
      text: "What is a data frame in R?",
      options: [
        "A type of graph",
        "A collection of vectors of equal length",
        "A single column of data",
        "A programming function"
      ]
    },
    {
      text: "How would you access the 'author' column in a data frame called 'books'?",
      options: [
        "`books[author]`",
        "`books$author`",
        "`author(books)`",
        "`books.author`"
      ]
    },
    {
      text: "What does `books[2, ]` return?",
      options: [
        "The second column of the 'books' data frame",
        "The second row of the 'books' data frame",
        "The second cell of the 'books' data frame",
        "An error"
      ]
    },
    {
      text: "In the expression `books[books$year > 1900, ]`, what does `books$year > 1900` do?",
      options: [
        "It selects all books published after 1900",
        "It creates a logical vector (TRUE/FALSE) for each row",
        "It returns an error because you can't compare a column to a number",
        "It modifies the 'year' column in the books data frame"
      ]
    },
    {
      text: "Which of the following would correctly filter the 'books' data frame to show only books by 'Austen'?",
      options: [
        "`books$author == 'Austen'`",
        "`books[books$author = 'Austen']`",
        "`books[books$author == 'Austen', ]`",
        "`filter(books, author == 'Austen')`"
      ]
    }
  ],
  answers: [1, 1, 1, 1, 2]
}, "quiz2");
</script>

## Hands-On Coding 💻

Let's apply our data frame skills to a new context: famous paintings. We'll create a data frame of artworks and practice accessing its data in various ways.

Try the following exercises:

1. Create a data frame called `paintings` with columns for title, artist, and year_created.

::: {.callout-note}
## Example Data

Feel free to use your own favorite paintings for this exercise! If you prefer, you can use the following information:

| Title                   | Artist            | Year Created |
|-------------------------|-------------------|--------------|
| Mona Lisa               | Leonardo da Vinci | 1503         |
| The Starry Night        | Vincent van Gogh  | 1889         |
| The Persistence of Memory | Salvador Dali   | 1931         |

:::

2. Access the 'artist' column of the data frame.
3. Get the third row of the data frame.
4. Find the year the second painting in the data frame was created.
5. (Optional challenge) Add a new column called 'style' to the data frame (e.g., "Renaissance", "Post-Impressionism", "Surrealism").


<div id="hidden-content2" style="display: none;">

```{r}
# 1. Create the paintings data frame
paintings <- data.frame(
    title = c("Mona Lisa", "The Starry Night", "The Persistence of Memory"),
    artist = c("Leonardo da Vinci", "Vincent van Gogh", "Salvador Dali"),
    year_created = c(1503, 1889, 1931)
)
# 2. Access the 'artist' column
paintings$artist
# 3. Get the third row
paintings[3, ]
# 4. Find the year the second painting was created
paintings[2, "year_created"]
# 5. Add a new 'style' column
paintings$style <- c("Renaissance", "Post-Impressionism", "Surrealism")
# View the updated data frame
print(paintings)
```


</div>

<script>
createToggleSection('button2', 'hidden-content2');
</script>

# Building Your Deck of Cards 🃏

Now that you have learned about different data types and how to work with data frames, let's put your skills to the test by creating a virtual deck of cards!

::: {.callout-tip collapse="false"}
### How to Use This Section {.unnumbered}

1. Read each task carefully.
2. Try solving it on your own first.
3. Use support tools if needed:
   - "Show Hint" 💡: For a gentle nudge 
   - "Show Template" 🧩: For a fill-in-the-blank version
   - "Show Solution" ✅: For the complete answer

   ::: {.callout-tip}
   ## Progressive Hints
   For some tasks, there are multiple hints available. For example, `Hint 1/3` means this the first out of three hints in total. Click on the `Show Hint` button will reveal one hint at a time.
   :::

### Tips for Success {.unnumbered}

- Type out the code yourself, even when using solutions.
- Experiment with small code changes to learn more.
- Take your time and review earlier sections if needed.

### Workflow Recommendations {.unnumbered}
[Explorers (for those who prefer more guidance)]{.highlight-green}:

1. Read the task carefully
2. Review the template for structure
3. Copy template to your file and fill in blanks
4. Use hints/solutions if needed, typing out the code
5. Experiment with small changes

[Adventurers (for those ready for more challenge)]{.highlight-blue}:

1. Attempt task independently
2. Use hints progressively if stuck
3. Compare your solution to the provided one
4. Try alternative approaches


### Remember {.unnumbered}

- Everyone starts as a beginner
- Mistakes are part of learning
- Focus on learning and improving, not perfection

Happy coding! 🚀
:::


## Create the Deck

::: {.callout-note}
## Structure of a Deck
Remember, a standard deck of cards consists of:

- 4 suits: Hearts♥️, Diamonds♦️, Clubs♣️, and Spades♠️
- 13 ranks in each suit: Ace, 2, 3, ..., 10, Jack, Queen, King
- A total of 52 cards (4 suits × 13 ranks)
:::

Create a data frame called `deck` with three columns: `suit`, `rank`, and `value`. 

- `suit` should contain all four suits, repeated 13 times each.
- `rank` should contain all 13 ranks, repeated for each suit.
- `value` should assign numeric values to the ranks (Ace = 1, Jack = 11, Queen = 12, King = 13, others as their numeric value).

<div id="hint1-1" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint 1/3
Recall how we created the `books` data frame. You can adapt it here. 

Instead of listing all cards one by one, you can use `rep()` function to create repeating patterns. Type `?rep()` in your console to learn how to use it. 
:::
</div>

<div id="hint1-2" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint 2/3
Use the `rep()` function to repeat values. For example, `rep(c("A", "B"), each = 3)` gives `"A" "A" "A" "B" "B" "B"`. rep(c("A", "B"), times = 3)` gives `"A" "B" "A" "B" "A" "B"`.

:::
</div>

<div id="hint1-3" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint 3/3
For the `suit` column, you'll need to repeat each suit 13 times. Try: 
`suit = rep(c("Hearts", "Diamonds", "Clubs", "Spades"), each = 13)`

For the `rank` column, you need to repeat the sequence of ranks 4 times. Use:
`rank = rep(c("Ace", 2:10, "Jack", "Queen", "King"), times = 4)`

For the `value` column, create a sequence from 1 to 13 and repeat it 4 times:
`value = rep(1:13, times = 4)`
:::
</div>

<script>
createProgressiveHints('hint1', 3);
</script>


<div id="template1" style="display: none;">

```{r, eval=FALSE}

deck <- data.frame(
    suit = rep(c("Hearts", "Diamonds", "Clubs", "Spades"), each = _),
    rank = rep(c("Ace", _, "Jack", "Queen", "King"), times = _),
    value = rep(_, times = 4)
)
```

</div>

<script>
createToggleSection('template-button1', 'template1', 'Show Template');
</script>

<div id="solution1" style="display: none;">


```{r}
deck <- data.frame(
suit = rep(c("Hearts", "Diamonds", "Clubs", "Spades"), each = 13),
rank = rep(c("Ace", 2:10, "Jack", "Queen", "King"), times = 4),
value = rep(c(1:13), times = 4)
)
# View the first few rows
head(deck)
```


</div>

<script>
createToggleSection('solution-button1', 'solution1');
</script>

## Accessing Information

Now that we have our deck, let's practice accessing information from it.

### How many cards are in the deck?
<div id="hint2-1-1" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint
To find the number of cards, think about how you'd check the structure of a data frame. 
:::
</div>

<script>
createProgressiveHints('hint2-1', 1);
</script>

<div id="template2-1" style="display: none;">

```{r, eval=FALSE}
# View the structure of the deck
_(deck)

# Count the number of cards (rows) in the deck
nrow(_)
```

</div>

<script>
createToggleSection('template-button2', 'template2-1', 'Show Template');
</script>

<div id="solution2" style="display: none;">
```{r}
str(deck)
nrow(deck)
```

::: {.callout-note}
## `nrow()` and Related Functions

`nrow()` is a useful function for working with data frames:

- `nrow(df)`: Returns the number of rows in a data frame
- `ncol(df)`: Returns the number of columns in a data frame
- `dim(df)`: Returns both the number of rows and columns as a vector
- `length(df)`: Returns the number of columns (as data frames are lists of vectors)

These functions are essential for understanding the size and structure of your data. For example, `nrow(deck)` tells you how many cards are in the deck.
:::

</div>

<script>
createToggleSection('solution-button2', 'solution2');
</script>

### What are all the unique suits in the deck?
<div id="hint2-2-1" style="display: none;">
::: {.callout-tip collapse="false"} 
## Hint
Check out the `unique()` function. This is useful when you want to see how many categories are in a factor or categorical variable.
:::
</div>

<div id="template2-2" style="display: none;">

```{r, eval=FALSE}
unique(deck$_)
```

</div>

<script>
createToggleSection('template-button2-2', 'template2-2', 'Show Template');
</script>

<script>
createProgressiveHints('hint2-2', 1);
</script>

<div id="solution2-2" style="display: none;">

```{r}
unique(deck$suit)
```

</div>

<script>
createToggleSection('solution-button2-2', 'solution2-2');
</script>

### How many hearts are in the deck?
<div id="hint2-3-1" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint
You can use logical indexing to select all hearts, then count them using `sum()`. How would you create a logical vector for hearts?
:::
</div>

<script>
createProgressiveHints('hint2-3', 1);
</script>

<div id="template2-3" style="display: none;">

```{r, eval=FALSE}
sum(deck$_ == "_")
```

</div>

<script>
createToggleSection('template-button2-3', 'template2-3', 'Show Template');
</script>

<div id="solution2-3" style="display: none;">

```{r}
sum(deck$suit == "Hearts")
```

</div>

<script>
createToggleSection('solution-button2-3', 'solution2-3');
</script>

### (Optional challenge) How to view all the face cards (Jack, Queen, King) in the deck?
<div id="hint2-5-1" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint
You need to do conditional filtering. You can use the `%in%` operator to check if values are in a set (i.e., multiple "OR" conditions). Check out the usage of this function from its help page.
:::
</div>

<script>
createProgressiveHints('hint2-5', 1);
</script>

<div id="template2-5" style="display: none;">

```{r, eval=FALSE}
deck[deck$rank _ c("_", "_", "_"), ]
```

</div>

<script>
createToggleSection('template-button2-5', 'template2-5', 'Show Template');
</script>

<div id="solution2-5" style="display: none;">

```{r}
deck[deck$rank %in% c("Jack", "Queen", "King"), ]
```

</div>

<script>
createToggleSection('solution-button2-5', 'solution2-5');
</script>

## Modifying the Deck

Let's make some changes to our deck:

### Add a face card indicator

Add a new column called `is_face_card` that is TRUE for Jack, Queen, and King, and FALSE for other cards.

<div id="hint3-1-1" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint
Think about how we can check if a card's rank is Jack, Queen, or King. Remember the `%in%` operator we used earlier.
:::
</div>

<script>
createProgressiveHints('hint3-1', 1);
</script>

<div id="template3-1" style="display: none;">

```{r, eval=FALSE}
deck$is_face_card <- deck$rank %in% c("_", "_", "_")
head(deck)
```

</div>

<script>
createToggleSection('template-button3-1', 'template3-1', 'Show Template');
</script>


<div id="solution3-1" style="display: none;">

```{r}
deck$is_face_card <- deck$rank %in% c("Jack", "Queen", "King")
head(deck)
```


</div>

<script>
createToggleSection('solution-button3-1', 'solution3-1');
</script>

### Create Jokers

Change the suit of all Aces to "Joker" (creating wild cards).

<div id="hint3-2-1" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint
Remember how we accessed specific rows in a data frame using `[,]`. How can we select all Aces and change their suit?
:::
</div>

<script>
createProgressiveHints('hint3-2', 1);
</script>

<div id="template3-2" style="display: none;">
```{r, eval=FALSE}
deck$_[deck$_ == "_"] <- "_"
deck[deck$_ == "_", ]
```
</div>

<script>
createToggleSection('template-button3-2', 'template3-2', 'Show Template');
</script>

<div id="solution3-2" style="display: none;">

```{r}
deck$suit[deck$rank == "Ace"] <- "Joker"
deck[deck$rank == "Ace", ]
```


</div>

<script>
createToggleSection('solution-button3-2', 'solution3-2');
</script>

### View high value cards

View only the cards with a value greater than 10.

<div id="hint3-3-1" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint
Recall how we used logical conditions to filter data frames. We can use different symbols such as `<` `>` `==`. 
:::
</div>

<script>
createProgressiveHints('hint3-3', 1);
</script>

<div id="template3-3" style="display: none;">

```{r, eval=FALSE}
high_value_cards <- deck[deck$_ _ _, ]
print(high_value_cards)
```

</div>

<script>
createToggleSection('template-button3-3', 'template3-3', 'Show Template');
</script>

<div id="solution3-3" style="display: none;">

```{r}
high_value_cards <- deck[deck$value > 10, ]
print(high_value_cards)
```


</div>

<script>
createToggleSection('solution-button3-3', 'solution3-3');
</script>

## Analyzing the Deck

Now that we've modified our deck, let's analyze it using the skills we've learned:

### Count face cards

Count how many face cards (Jack, Queen, King) are in the deck.

<div id="hint4-1-1" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint
Remember the `is_face_card` column we created? How can we count the number of TRUE values? Try `sum()`.
:::
</div>

<script>
createProgressiveHints('hint4-1', 1);
</script>

<div id="template4-1" style="display: none;">

```{r, eval=FALSE}
face_card_count <- sum(deck$_)
print(paste("Number of face cards:", _))
```

</div>

<script>
createToggleSection('template-button4-1', 'template4-1', 'Show Template');
</script>

<div id="solution4-1-1" style="display: none;">

```{r}
face_card_count <- sum(deck$is_face_card)
print(paste("Number of face cards:", face_card_count))
```

</div>

<script>
createToggleSection('solution-button4-1', 'solution4-1-1');
</script>

### Find average card value

Calculate the average value of all cards in the deck.

<div id="hint4-2-1" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint
We can use `mean()` to calculate an average. Which column contains the card values?
:::
</div>

<script>
createProgressiveHints('hint4-2', 1);
</script>

<div id="template4-2" style="display: none;">

```{r, eval=FALSE}
average_value <- _(deck$_)
print(paste("Average card value:", round(_, 2)))
```

</div>

<script>
createToggleSection('template-button4-2', 'template4-2', 'Show Template');
</script>

<div id="solution4-2" style="display: none;">

```{r}
average_value <- mean(deck$value)
print(paste("Average card value:", round(average_value, 2)))
```

</div>

<script>
createToggleSection('solution-button4-2', 'solution4-2');
</script>

### Find highest value card

Identify the card(s) with the highest value in the deck.

<div id="hint4-3-1" style="display: none;">
::: {.callout-tip collapse="false"}
## Hint
How can we find the maximum value in the deck? Try `max()`.

Once we have that, how can we select all cards with that value?
:::
</div>

<script>
createProgressiveHints('hint4-3', 1);
</script>

<div id="template4-3" style="display: none;">

```{r, eval=FALSE}
max_value <- _(deck$value)
highest_cards <- deck[deck$value == _, ]
print(_)
```

</div>

<script>
createToggleSection('template-button4-3', 'template4-3', 'Show Template');
</script>

<div id="solution4-3" style="display: none;">

```{r}
max_value <- max(deck$value)
highest_cards <- deck[deck$value == max_value, ]
print(highest_cards)
```



</div>

<script>
createToggleSection('solution-button4-3', 'solution4-3');
</script>


# Conclusion

::: {.callout-note}
## Key Takeaways
In this chapter, we've covered:

- Creating and manipulating data frames in R
- Understanding different R data types
- Building a virtual deck of cards
- Accessing and modifying data within data frames
- Performing simple analyses on our card deck
:::

These skills form an essential foundation for working with structured data in R. As we progress in our digital humanities journey, we'll build upon these concepts to perform more complex data manipulations and analyses.

::: {.topic-flow}
::: {.topic-box .highlight-blue}
Data Types
:::
::: {.topic-box .highlight-green}
Data Frames
:::
::: {.topic-box .highlight-yellow}
Data Access
:::
::: {.topic-box .highlight-pink}
Data Analysis
:::
:::

<iframe src="float_traffic.html" width="200px" height="200px" style="border:none; position: fixed; bottom: 10px; right: 10px; z-index: 9999;" scrolling="no"></iframe>


<iframe src="float_dash.html" width="200px" height="200px" style="border:none; position: fixed; bottom: 10px; left: 10px; z-index: 9999;" scrolling="no"></iframe>

